Jane, John … Leslie? A Historical Method for Algorithmic Gender Detection

### 2014-09-09 article.Rmd
Finished description of data

TODO IPUMS

>>>>

The IPUMS-USA dataset, contributed by Benjamin Schmidt, contains records from the United States decennial census from 1790 to 1930. This dataset contains the birth year and numbers of males and females under the age of 62 for all the years in that range. This data has been aggregated by IPUMS at the University of Minnesota and is released as a sample of the total census data. Unlike the SSA dataset, which includes 100% of every name reported to the Social Security Administration, the IPUMS data contains 5% or 10% samples. Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons.^[Steven Ruggles, J. Trent Alexander, Katie Genadek, Ronald Goeken, Matthew B. Schroeder, and Matthew Sobek. Integrated Public Use Microdata Series: Version 5.0 [Machine-readable database]. Minneapolis, MN: Minnesota Population Center [producer and distributor], 2010.]

These two datasets (and any future datasets to be added to the package) thus contain a simple table of data. The columns

^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons. By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster. Hadley Wickham, "[Tidy Data](http://vita.had.co.nz/papers/tidy-data.html)," *Journal of Statistical Software* (forthcoming).] The columns contain the `name`, `year`, and number of `female` and `male` uses of a name. Each row contains a combination for a name and a year.

```{r ssa_national, echo=TRUE, results='markup'}
genderdata::ssa_national
```

```{r ipums_usa, echo=TRUE, results='markup'}
genderdata::ipums_usa
```
